A hierarchical model to evaluate pest treatments from prevalence and intensity data

### 2023-09-03 Article.qmd
Modifications to generate tables 4, 5, 6 and 7 directly from the ouput, using kbl from kableExtra - Warning : I didn't manage to get the % in the captions (table 4 and 5 are thus missing % for 100% Y, ..., need to be fixed)

keywords: [bayesian model, epidemiology, hierarchical model, pest control, trial, survey]
doi: "to complete"
citation:
  type: article-journal
  container-title: "Computo"
  issued: "2023/01/25"
  url: https://computo.sfds.asso.fr/
  issn: "2824-7795"
google-scholar: true
bibliography: references.bib
github-user: computorg
repo: "To complete"
draft: false 
published: false
format:
  computo-html: default
  computo-pdf: default
---

```{r dependencies, cache = FALSE, include = FALSE}
library(ggplot2);
library(tidyverse);
library(gridExtra);
library(rjags);
library(latex2exp);
library(grid);
library(dplyr);
library(tidyr);
library(tibble)

@@

```{r simulations, cache = TRUE, eval = TRUE, results = 'hide'}
J = 4;   T = 3;   N = 10;   K = 4;

@@

|**Parameter** | **Mean** | **SD** | **2.5%** | **97.5%**|
|---------|:-----|:-----|:-----|:-----|
|$\alpha_0$ | 0.50 | 0.34 | -0.18 | 1.18 |
|$\alpha_1$ | 0.16 | 0.06 | 0.04 | 0.29 |
|$\sigma_0$ | 1.87 | 0.26 | 1.44 | 2.45 |
|$\chi$ | 0.27 | 0.06 | 0.15 | 0.40 |
|$\eta$ | 0.98 | 0.03 | 0.92 | 1.03 |
|$\gamma_0$ - untreated | 0.00 | 0.00 | 0.00 | 0.00 |
|$\gamma_0$ - Mavrik Jet | -0.13 | 0.16 | -0.45 | 0.17 |
|$\gamma_0$ - Movento | -1.13 | 0.12 | -1.37 | -0.89 |
|$\gamma_0$ - Teppeki | -1.24 | 0.11 | -1.46 | -1.03 |
|$\gamma_1$ - untreated | 0.00 | 0.00 | 0.00 | 0.00 |
|$\gamma_1$ - Mavrik Jet | 0.24 | 0.13 | -0.01 | 0.48 |
|$\gamma_1$ - Movento | -0.14 | 0.10 | -0.34 | 0.05 |
|$\gamma_1$ - Teppeki | -0.15 | 0.09 | -0.32 | 0.02|

: Summary of the posterior distributions obtained with the "100% Y" scenario for the model parameters and treatment efficacies: posterior mean, standard deviation, 2.5% and 97.5% quantiles.

@@

| Insecticide | Data | At 6 days | At 12 days |
|---------|:-----|:-----|:-----|
| Mavrik Jet | 100% Y | - 27.7 | - 41.3 |
| Movento | 100% Y | -  40.5 | - 46.7 |
| Teppeki | 100% Y | - 25.6 | - 37.8 |
| Mavrik Jet | 50% Y / 50% Z | - 18.4 | - 31.1 |
| Movento | 50% Y / 50% Z | - 27.1 | - 31.4 |
| Teppeki | 50% Y / 50% Z | - 12.1 | - 22.4 |
| Mavrik Jet | 50% Z | 26.7  | - 19.1  |
| Movento | 50% Z | 0.8  | - 9.8 |
| Teppeki | 50% Z | 44.4 | 16.9 |

: Differences in the sizes of the 95% credibility intervals (CI) of the estimated treatment efficacies for the scenarios "50% Y / 50% Z", "50% Z" and "100% Y", compared to "50% Y". The difference is given in percentage. A positive (negative) value indicates an increase (decrease) of the credibility interval size. The third column indicates differences for the efficacy at 6 days after pesticide spray, and the fourth column indicates the difference for the efficacy at 12 days.

@@

|$\alpha_0$ | $\alpha_1$ | $\gamma_{01}$ | $\gamma_{02}$ | $\gamma_{03}$ | $\gamma_{04}$ | $\gamma_{11}$ | $\gamma_{12}$ | $\gamma_{13}$ | $\gamma_{14}$ | $\sigma_0$ | $\eta$ | $\chi$|
|---------|:-----|:-----|:-----|:-----|:-----|:-----|:-----|:-----|:-----|:-----|:-----|:-----|
|$\text{\small 0.5}$ |$\text{\small 0.16}$ |$\text{\small 0}$ |$\text{\small -0.13}$ |$\text{\small -1.13}$ |$\text{\small -1.24}$ |$\text{\small 0}$ |$\text{\small 0.24}$ |$\text{\small -0.14}$ |$\text{\small -0.15}$ |$\text{\small 1.87}$ |$\text{\small 0.27}$ |$\text{\small 0.98}$|

@@

| Set | $\ \alpha_0$ | $\alpha_1$ | $\gamma_{01}$ | $\gamma_{02}$ | $\gamma_{03}$ | $\gamma_{04}$ | $\gamma_{11}$ | $\gamma_{12}$ | $\gamma_{13}$ | $\gamma_{14}$ | $\sigma_0$ | $\eta$ | $\chi$|
|-|---------|:-----|:-----|:-----|:-----|:-----|:-----|:-----|:-----|:-----|:-----|:-----|:-----|
| $\text{\small A}$ | $\textbf{\small\ -1}$ | $\text{\small 0.16}$ | $\text{\small 0}$ | $\text{\small -0.13}$ | $\text{\small -1.13}$ | $\text{\small -1.24}$ | $\text{\small 0}$ | $\text{\small 0.24}$ | $\text{\small -0.14}$ | $\text{\small -0.15}$ | $\text{\small 0.3}$ | $\text{\small 0.27}$ | $\text{\small 0.98}$|
| $\text{\small B}$ | $\textbf{\small\ 1}$ | $\text{\small 0.16}$ | $\text{\small 0}$ | $\text{\small -0.13}$ | $\text{\small -1.13}$ | $\text{\small -1.24}$ | $\text{\small 0}$ | $\text{\small 0.24}$ | $\text{\small -0.14}$ | $\text{\small -0.15}$ | $\text{\small 0.3}$ | $\text{\small 0.27}$ | $\text{\small 0.98}$ |
| $\text{\small C}$ | $\textbf{\small\ 2}$ | $\text{\small 0.16}$ | $\text{\small 0}$ | $\text{\small -0.13}$ | $\text{\small -1.13}$ | $\text{\small -1.24}$ | $\text{\small 0}$ | $\text{\small 0.24}$ | $\text{\small -0.14}$ | $\text{\small -0.15}$ | $\text{\small 0.3}$ | $\text{\small 0.27}$ | $\text{\small 0.98}$|

: Parameters considered for the design of future pest surveys.

>>>>

keywords: [bayesian model, epidemiology, hierarchical model, pest control, trial, survey]
doi: "to complete"
citation:
  type: article-journal
  container-title: "Computo"
  issued: "2023/01/25"
  url: https://computo.sfds.asso.fr/
  issn: "2824-7795"
google-scholar: true
bibliography: references.bib
github-user: computorg
repo: "To complete"
draft: false 
published: false
format:
  computo-pdf: default
---

```{r dependencies, cache = FALSE, include = FALSE}
library(ggplot2);
library(tidyverse);
library(gridExtra);
library(rjags);
library(latex2exp);
library(grid);
library(dplyr);
library(tidyr);
library(tibble)
library(kableExtra)

@@

```{r simulations, cache = FALSE, eval = TRUE, results = 'hide'}
J = 4;   T = 3;   N = 10;   K = 4;

@@

```{r table4}
#| echo: false
tab4 = cbind(summary(resY)$statistics, summary(resY)$quantiles) %>% 
    as.data.frame %>%
    rownames_to_column(var = "Parameter") %>% select(- c(4, 5, c(7 : 9))) %>%
    mutate(Parameter = recode(Parameter, "gamma0[1]" = "gamma0 - untreated", "gamma0[2]" = "gamma0 - Mavrik Jet",
                                         "gamma0[3]" = "gamma0 - Movento", "gamma0[4]" = "gamma0 - Teppeki",
                                          "gamma1[1]" = "gamma1 - untreated", "gamma1[2]" = "gamma1 - Mavrik Jet",
                                         "gamma1[3]" = "gamma1 - Movento", "gamma1[4]" = "gamma1 - Teppeki"  )) %>%
    filter(!grepl("Eff", Parameter)) %>%

    mutate(Parameter = recode(Parameter, "alpha0" = "\u03B1\u2080",
                                         "alpha1" = "\u03B1\u2081",
                                         "chi" = "\u03C7",
                                         "eta" = "\u03B7",
                                         "sigma0" = "\u03C3\u2080")) %>%
    mutate(Parameter = gsub(x = Parameter, pattern = "gamma0", replacement = "\u03B3\u2080")) %>%
    mutate(Parameter = gsub(x = Parameter, pattern = "gamma1", replacement = "\u03B3\u2081")) %>%
    mutate_if(is.numeric, round, 2)

# re-ordering the parameters
tab4 = tab4[c(c(1 : 2), 13, c(3 : 12)), ]

rownames(tab4) = NULL

tab4 %>% 
  kbl(booktabs = TRUE, caption = "Summary of the posterior distributions obtained with the 100 Y scenario for the model parameters and treatment efficacies: posterior mean, standard deviation, 2.5 and 97.5 quantiles.") %>%
  kable_styling()
```

@@

```{r table5}
#|echo: false
cbind(df_len6, df_len12 %>% select(- Data, - Insecticide)) %>% 
  kbl(booktabs = TRUE, caption = "Differences in the sizes of the 95 credibility intervals (CI) of the estimated treatment efficacies for the scenarios 50 Y / 50 Z, 50 Z and 100 Y, compared to 50 Y. The difference is given in percentage. A positive (negative) value indicates an increase (decrease) of the credibility interval size. The third column indicates differences for the efficacy at 6 days after pesticide spray, and the fourth column indicates the difference for the efficacy at 12 days.") %>% 
  kable_styling()
```

@@

```{r table6}
#|echo: false

tab6 = data.frame(Parameters = c(

                    # alpha_0 et alpha_1
                    c("\u03B1\u2080", "\u03B1\u2081"),

                    # gamma_0
                    c("\u03B3\u2080\u2080", "\u03B3\u2080\u2081", "\u03B3\u2080\u2082", "\u03B3\u2080\u2083"),

                    # gamma_1
                    c("\u03B3\u2081\u2080", "\u03B3\u2081\u2081", "\u03B3\u2081\u2082", "\u03B3\u2081\u2083"),

                    # sigma_0, eta et chi
                    c("\u03C3\u2080", "\u03B7", "\u03C7")
                    ),
                  Values = c(alpha0, alpha1, gamma0, gamma1, sig0, eta, chi))

tab6 %>%
    t %>%
    kbl(booktabs = TRUE, caption = "Parameters used to generate virtual data.") %>%
    kable_styling()
```

@@

```{r table7}
#|echo: false

tab7 = data.frame(Parameters = c(
                    # alpha_0 et alpha_1
                    c("\u03B1\u2080", "\u03B1\u2081"),

                    # gamma_0
                    c("\u03B3\u2080\u2080", "\u03B3\u2080\u2081", "\u03B3\u2080\u2082", "\u03B3\u2080\u2083"),

                    # gamma_1
                    c("\u03B3\u2081\u2080", "\u03B3\u2081\u2081", "\u03B3\u2081\u2082", "\u03B3\u2081\u2083"),

                    # sigma_0, eta et chi
                    c("\u03C3\u2080", "\u03B7", "\u03C7")
                    ),

                  "Values set A" = c(-1, alpha1, gamma0, gamma1, sig0, eta, chi),
                  "Values set B" = c(1, alpha1, gamma0, gamma1, sig0, eta, chi),
                  "Values set C" = c(2, alpha1, gamma0, gamma1, sig0, eta, chi)
                  )

tab7 %>%
    t %>% 
    `rownames<-`(c("Set", "A", "B", "C")) %>% 
    kbl(booktabs = TRUE, caption = "Parameters considered for the design of future pest surveys.") %>%
    column_spec(2, bold = TRUE) %>%
    kable_styling()
```
